The Poisson approximate likelihood compared to the particle filter

### 2024-07-25 ms.Rnw
some tidying

\documentclass[10pt]{article}
\usepackage{color}
\usepackage{graphicx} % Required for inserting 
%%\usepackage{algpseudocode} % Include the package for algorithmic environment
\usepackage{algorithm}
\usepackage{booktabs} % For better-quality horizontal lines
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{threeparttable} % For table notes
\usepackage{array} % For table formatting

\usepackage{setspace}
\usepackage{float} % add this in your preamble
\usepackage{amsmath}
\usepackage{tabularx}

@@

\usepackage[round]{natbib}

\usepackage{multirow}
\usepackage{fourier} 
\usepackage{booktabs}
\usepackage{array}
\usepackage{makecell}
\usepackage[font={small,it}]{caption}
\usepackage[utf8]{inputenc}
\usepackage{mathtools}
\usepackage{subcaption}
\usepackage{tabstackengine}
\usepackage[toc,page]{appendix}
\usepackage{graphicx, fullpage, verbatim, amsmath}
\usepackage{url, amsfonts, amssymb, amsthm,color, enumerate}
\usepackage{placeins, listings, textcomp, mathtools, multicol, tikz}
\usepackage{dsfont}
\usepackage{amsfonts}
\usepackage{bm}

\TABstackMath

\doublespacing
% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

@@

\begin{document}
%\SweaveOpts{concordance=TRUE}

@@

library("knitr")
opts_knit$set(concordance=TRUE)
opts_chunk$set(
    concordance = TRUE,
    tidy = TRUE,
    message = FALSE,
    warning = FALSE,
    tidy.opts = list(
        keep.blank.line = FALSE
    ),
    comment = "",
    echo = FALSE,
    results = FALSE,
    dev.args = list(
        bg = "transparent",
        pointsize = 9
    ),
    fig.path = "figure/"
)

@@

<<OvOv-model, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
# measurement model 
dmeas <- Csnippet("
                  if (ISNA(cases1)) {
                  lik = (give_log) ? 0 : 1;
                  } else {
                        lik =  dbinom(cases1, H1, q1, 1) +
                        dbinom(cases2, H2, q2, 1) +
                        dbinom(cases3, H3, q3, 1);

@@

<<maximization1, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
set.seed(123)
sir_box <- list(
  beta1=c(10,15),
  beta2=c(0.2,0.4),
  beta3=c(0.3,0.5),
  phi=c(0.01,0.3),
  beta11=c(0.1,0.2),
  sigma_q=c(0.001,0.1),
  sigma_xi=c(65,70),
  S10 = c(0.01,0.05),
  I10 = c(0.0001, 0.0005),
  R10 = c(0.01, 0.05),
  S20 = c(0.01,0.7),
  I20 = c(0.000001,0.00002),
  R20 = c(0.001,0.005),
  S30 = c(0.01,0.5),
  I30 = c(0.000001, 0.0001),
  R30 = c(0.0001,0.005),
  #fixed params
  gamma = c(1,1),
  delta1=c(1/(5*52),1/(5*52)),
  delta2=c(1/(55*52),1/(55*52)),
  alpha=c(1/(78.86912*52),1/(78.86912*52)),
  mu=c(0,0),
  N=c(82372825,82372825),
  omega=c(1/(1*52),1/(1*52))
)

@@

<<maximization2, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
set.seed(1234)
sd_02 = 2/3

@@

<<maximization3, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
set.seed(12345)
sd_03           = (2/3)^2

@@

<<estimated-mle, echo=FALSE, message=FALSE, warning=FALSE>>=
mif_el <- readRDS(paste0(out_dir,"round_03/ovov_el_03.rds")) 
# Can be changed if more rounds needed
est_mle_vec <- mif_el[which.max(mif_el$logLik),]
@

@@

<<pf-sim-100, echo=FALSE, results='hide', message=FALSE, warning=FALSE>>=
# Set WWR's MLE
wwr_mle_raw <- c(params_fixed, 
             "beta1" = 11.48,
             "beta2" = 0.25,
             "beta3" = 0.35,
             "phi" = 0.14,
             "beta11" = 0.16,
             "sigma_q" = 0.021,
             "sigma_xi" = 66.89)

@@

<<pfilter_sim_100, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
ovov_sim_100 <- readRDS(paste0(out_dir, "pfilter/ovov_sim_100.rds"))
bake(file=paste0(out_dir, "pfilter/pfilter_ovov_sim_100.rds"),{
  pfilter_ovov_sim_100 <- c()
  registerDoFuture()
  plan(multicore, workers=ncores)

@@

<<palfilter_sim_100, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
ovov_sim_100 <- readRDS(paste0(out_dir, "pfilter/ovov_sim_100.rds"))
prop <- 420

>>>>

\documentclass[10pt]{article}
\usepackage{color}
\usepackage{graphicx} % Required for inserting
\usepackage{booktabs} % For better-quality horizontal lines
\usepackage{multirow}
\usepackage{threeparttable} % For table notes
\usepackage{array} % For table formatting
\usepackage{setspace}
\usepackage{float} % add this in your preamble
\usepackage{amsmath}
\usepackage{tabularx}

@@

\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{array}
\usepackage{makecell}
\usepackage[font={small,it}]{caption}
\usepackage[utf8]{inputenc}
\usepackage{mathtools}
\usepackage{subcaption}
\usepackage{graphicx, fullpage, verbatim, amsmath}
\usepackage{url, amsfonts, amssymb, amsthm,color, enumerate}
\usepackage{amsfonts}
\usepackage{bm}

\doublespacing
% \usepackage[english]{babel}

@@

\begin{document}

@@

library("knitr")
opts_knit$set(concordance=TRUE)
opts_chunk$set(
#    concordance = TRUE,
    tidy = TRUE,
#    message = FALSE,
#    warning = FALSE,
    tidy.opts = list(
        keep.blank.line = FALSE
    ),
    comment = "",
    echo = FALSE,
    results = FALSE,
    dev.args = list(
        bg = "transparent",
        pointsize = 9
    ),
    fig.path = "figure/"
)

@@

<<OvOv-model>>=
# measurement model 
dmeas <- Csnippet("
                  if (ISNA(cases1)) {
                  lik = (give_log) ? 0 : 1;
                  } else {
                        lik =  dbinom(cases1, H1, q1, 1) +
                        dbinom(cases2, H2, q2, 1) +
                        dbinom(cases3, H3, q3, 1);

@@

<<maximization1>>=
set.seed(123)
sir_box <- list(
  beta1=c(10,15),
  beta2=c(0.2,0.4),
  beta3=c(0.3,0.5),
  phi=c(0.01,0.3),
  beta11=c(0.1,0.2),
  sigma_q=c(0.001,0.1),
  sigma_xi=c(65,70),
  S10 = c(0.01,0.05),
  I10 = c(0.0001, 0.0005),
  R10 = c(0.01, 0.05),
  S20 = c(0.01,0.7),
  I20 = c(0.000001,0.00002),
  R20 = c(0.001,0.005),
  S30 = c(0.01,0.5),
  I30 = c(0.000001, 0.0001),
  R30 = c(0.0001,0.005),
  #fixed params
  gamma = c(1,1),
  delta1=c(1/(5*52),1/(5*52)),
  delta2=c(1/(55*52),1/(55*52)),
  alpha=c(1/(78.86912*52),1/(78.86912*52)),
  mu=c(0,0),
  N=c(82372825,82372825),
  omega=c(1/(1*52),1/(1*52))
)

@@

<<maximization2>>=
set.seed(1234)
sd_02 = 2/3

@@

<<maximization3>>=
set.seed(12345)
sd_03           = (2/3)^2

@@

<<estimated-mle>>=
mif_el <- readRDS(paste0(out_dir,"round_03/ovov_el_03.rds")) 
# Can be changed if more rounds needed
est_mle_vec <- mif_el[which.max(mif_el$logLik),]
@

@@

<<pf-sim-100>>=
# Set WWR's MLE
wwr_mle_raw <- c(params_fixed, 
             "beta1" = 11.48,
             "beta2" = 0.25,
             "beta3" = 0.35,
             "phi" = 0.14,
             "beta11" = 0.16,
             "sigma_q" = 0.021,
             "sigma_xi" = 66.89)

@@

<<pfilter_sim_100>>=
ovov_sim_100 <- readRDS(paste0(out_dir, "pfilter/ovov_sim_100.rds"))
bake(file=paste0(out_dir, "pfilter/pfilter_ovov_sim_100.rds"),{
  pfilter_ovov_sim_100 <- c()
  registerDoFuture()
  plan(multicore, workers=ncores)

@@

<<palfilter_sim_100>>=
ovov_sim_100 <- readRDS(paste0(out_dir, "pfilter/ovov_sim_100.rds"))
prop <- 420
